### 2024-10-20 src/Submission.Rnw
weird edit

\begin{document}

@@

\section{Maximum Likelihood Estimation (MLE)}

\section{Likelihood Ratio Test}

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

\section{Maximum Likelihood Estimation (MLE)}
<<echo=TRUE>>=
diabetes <- read.csv("Group Project/dat/diabetes.csv", header = TRUE)
head(diabetes)

neg_log_likelihood <- function(params, data) {
  # Extract the parameters
  beta0 <- params[1]        # Intercept
  beta1 <- params[2]        # Slope
  sigma <- params[3]        # Standard Deviation

  # Extract dependent and independent variables
  x <- data$bmi
  y <- data$target

  # Predicted value of a linear model
  y_pred <- beta0 + beta1 * x

  # Log-likelihood for normal distribution
  norm_log_likelihood <- -sum(dnorm(y, mean = y_pred, sd = sigma, log = TRUE))

  return(norm_log_likelihood)
}
# Now to use optim() to minimise the negative log-likelihood and estimate params
int_params <- c(beta0 = 0, beta1 = 0, sigma = 1)

mle <- optim(
  par = int_params,           # Initial guess
  fn = neg_log_likelihood,    # The function being minimised
  data = diabetes,            # Data to be used
  method = "BFGS",            # Optimisation method (less fragile and more common)
  hessian = TRUE              # Return Hessian for variance estimation
)

# Results
mle$par

# Predicted values from the model put back into likelihood function for LP
y_pred <- mle$par[1] + mle$par[2] * diabetes$bmi

# Residuals
residuals <- diabetes$target - y_pred

# Plot residuals to check for patterns
plot(diabetes$bmi, residuals, main = "Residuals vs BMI", xlab = "BMI", ylab = "Residuals")
@

\section{Likelihood Ratio Test}
<<echo=TRUE>>=
# Fitting Model 1: Progression ~ BMI
model1 <- lm(target ~ bmi, data = diabetes)

# Fitting Model 2: Progression ~ BMI + Age
model2 <- lm(target ~ bmi + age, data = diabetes)

# Summarise the models
summary(model1)
summary(model2)

# Extract the log-likelihood for both models
logLik_model1 <- logLik(model1)
logLik_model2 <- logLik(model2)

# Print the log-likelihoods (optional)
logLik_model1
logLik_model2

# Calculate the likelihood ratio test statistic
LRT_stat <- 2 * (logLik_model2 - logLik_model1)

# Degrees of freedom is the difference in the number of parameters
df <- df.residual(model1) - df.residual(model2)

# Compute the p-value for the test
p_value <- pchisq(LRT_stat, df = df, lower.tail = FALSE)

# Print the result
cat("Likelihood Ratio Test Statistic:", LRT_stat, "\n")
cat("p-value:", p_value, "\n")
@
